Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-04-10 WSSSPE2_report/WSSSPE2_review.tex
Section 3.2: fix typos.

Neil Chue Hong is the Director of the Software Sustainability Institute (SSI) in the 
United Kingdom. The SSI was founded to support the UK?s research software 
community by cultivating better, more sustainable, research software to enable 
world-class research. Chue Hong's keynote\footnote{Slides available at: http://dx.doi.org/10.6084/m9.figshare.1243288}
 started by making the point of that the use of 
- and reliance on - software is pervasive in all areas of world-leading research, 
showing examples from disciplines as diverse as the humanities and high-energy 
physics, and quoting Kersten Kleese Van Dam of the Pacific Northwestern 
National Laboratory via a petition campaign\footnote{https://www.change.org/p/everyone-in-the-research-community-we-must-accept-that-software-is-fundamental-to-research-or-we-will-lose-our-ability-to-make-groundbreaking-discoveries}
at change.org, ``\emph{Today there are very few science areas left 
which do not rely on IT and thus software for the majority of their research work. 
More importantly key scientific advances in experimental and observational science 
would have been impossible without better software.}'' He also cited Daniel Katz, 
Software Infrastructure for Sustained Innovation(SI2) Program Director of the 
National Science Foundation, ``\emph{Scientific discovery and innovation are 
advancing along fundamentally new pathways opened by development of 
increasingly sophisticated software. Software is an integral enabler of 
computation, experiment and theory, and directly responsible for increased 
scientific productivity and enhancement of researchers' capabilities.}''

Chue Hong drew attention to the issue that in the cyber-infrastructure and high 
performance community, we all too often disregard the hundreds of thousands 
of researchers developing software thinking they are the long tail. In fact, the 
numbers point to the fact that they are actually the mainstream and we (referring 
to people including the audience of the talk) are the elite minority. He
emphasized that software is no longer special; it is both essential to and
common in scientific research. A 2014 survey\footnote{http://www.software.ac.uk/blog/2014-12-04-its-impossible-conduct-research-without-software-say-7-out-10-uk-researchers} 
conducted by the SSI polled researchers 
from 15 research-intensive UK universities (406 respondents covering a 
representative range of funders, disciplines, and seniority) reported that 92\% 
confirmed the use of research software  and 89\% affirmed that it would be 
impossible or difficult to conduct research without software. Nevertheless the British
research community is just starting to understand the magnitude of the issue. 
Whilst many researchers make use of software such as MATLAB, SPSS, and 
Excel, data from the aforementioned SSI survey shows that over half (56\%) of 
respondents developed their own research software (which equates to over 
140,000 researchers if extrapolated across the UK) and yet 71\% had no formal 
software-development training, having to rely on their own coding skills.

>>>>

Neil Chue Hong is the Director of the Software Sustainability Institute (SSI) in the 
United Kingdom. The SSI was founded to support the UK's research software 
community by cultivating better, more sustainable research software to enable 
world-class research. Chue Hong's keynote\footnote{Slides available at: http://dx.doi.org/10.6084/m9.figshare.1243288}
 started by making the point of that the use of 
-- and reliance on -- software is pervasive in all areas of world-leading research, 
showing examples from disciplines as diverse as the humanities and high-energy 
physics, and quoting Kersten Kleese Van Dam of the Pacific Northwestern 
National Laboratory via a petition campaign\footnote{https://www.change.org/p/everyone-in-the-research-community-we-must-accept-that-software-is-fundamental-to-research-or-we-will-lose-our-ability-to-make-groundbreaking-discoveries}
at change.org, ``\emph{Today there are very few science areas left 
which do not rely on IT and thus software for the majority of their research work. 
More importantly key scientific advances in experimental and observational science 
would have been impossible without better software.}'' He also cited Daniel Katz, 
Software Infrastructure for Sustained Innovation(SI2) Program Director of the 
National Science Foundation, ``\emph{Scientific discovery and innovation are 
advancing along fundamentally new pathways opened by development of 
increasingly sophisticated software. Software is an integral enabler of 
computation, experiment and theory, and directly responsible for increased 
scientific productivity and enhancement of researchers' capabilities.}''

Chue Hong drew attention to the issue that in the cyber-infrastructure and 
high-performance community, we all too often disregard the hundreds of thousands 
of researchers developing software thinking they are the long tail. In fact, the 
numbers point to the fact that they are actually the mainstream and we (referring 
to people including the audience of the talk) are the elite minority. He
emphasized that software is no longer special; it is both essential to and
common in scientific research. A 2014 survey\footnote{http://www.software.ac.uk/blog/2014-12-04-its-impossible-conduct-research-without-software-say-7-out-10-uk-researchers} 
conducted by the SSI polled researchers 
from 15 research-intensive UK universities (406 respondents covering a 
representative range of funders, disciplines, and seniority) reported that 92\% 
confirmed the use of research software  and 89\% affirmed that it would be 
impossible or difficult to conduct research without software. Nevertheless the British
research community is just starting to understand the magnitude of the issue. 
Whilst many researchers make use of software such as MATLAB, SPSS, and 
Excel, data from the aforementioned SSI survey shows that over half (56\%) of 
respondents developed their own research software (which equates to over 
140,000 researchers if extrapolated across the UK) and yet 71\% had no formal 
software-development training, having to rely on their own coding skills.
